Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-03 qhtv2.tex
Allègement des notations

\paragraph{Duplicate detection problem.}
We consider the following problem (duplicate detection problem, DPP): given a stream $S_t = \{s_t, s_{t-1}, \dotsc\}$ and an item $s^\star$, find whether $s^\star \in S_t$. It is understood that the stream $S_t$ is updated at every time increment. An \emph{efficient} solution to this problem provides an answer with bounded error $\epsilon$, using polynomially many resources (time and memory) in $1/epsilon$.

@@

\begin{definition}[Duplicate and unseen element]
Let $E = \{e_i\}$ be a stream over $\Gamma$. Some element $e_j \in E$ is called a \emph{duplicate} if and only if there exists $i < j$ such that $e_i = e_j$. Otherwise, $e_j$ is said to be \emph{unseen}.
\end{definition}

@@

\begin{definition}[False positive, false negative]
	Let $e \in E$ be an unseen (resp. duplicate) and $\mathcal F$ be a filter. We say that $e$ is a \emph{false positive} (resp. \emph{false negative}) when $\mathcal F.\mathsf{Detect}(e) = \DUPLICATE$ (resp. $\UNSEEN$).
\end{definition}

@@

\subsection{Filter Saturation}
While the previous definition sounds interesting and is often simpler to work with, especially for the derivation of the FPR and FNR, it has been shown that this kind of duplicate detection is an ill-formed problem, as filters saturate and reach the asymptotic relation $\FPR_\infty + \FNR_\infty = 1$, which means that the filter is not any more efficient than a filter answering randomly. We reproduce the statement and proof from \cite{10.1145/3297280.3297335}.

	\begin{theorem}[\cite{10.1145/3297280.3297335}]\label{thm:sat_u}
	Let $\FNR_\infty$ and $\FPR_\infty$ be the asymptotic false negative and false positive rates of a filter of $M$ bits of memory (i.e. the FNR and FPR on a stream of size going to infinity).

	If $U \gg M$, then $\FNR_\infty + \FPR_\infty = 1$, which characterises random filters (i.e. filters answering randomly to any query).
\end{theorem}

\begin{proof}
	First note that random filters always verify the relation $\FNR + \FPR = 1$: given that a random filter will return \DUPLICATE with a probability of $p$, an unseen element will be classified as \DUPLICATE with probability $p$, and a duplicate will be classified as \UNSEEN with probability $1-p$, hence the result.

	Now, on infinite streams with infinitely many different elements, filters are saturated with information. Given that a filter can only store at most one element per bit (cf. \cite[][Section 2.1]{10.1145/3297280.3297335}), thus a filter of size $M$ can remember at most $M$ elements. However, after $fM$ insertions (with $f \gg 1$), the filter remembers at most a tiny fraction $\frac 1f \ll 1$ of the stream. Having reached its saturated state, the filter has an extremely tiny probability $\frac 1f \approx 0$ of correctly guessing whether the incoming element is a duplicate or not, given what the filter actually knows about the stream. For this reason, the best strategy of a saturated filter is almost indistinguishable from a random strategy, i.e. randomly outputting \DUPLICATE. When the stream grows indefinitely, the filter becomes asymptotically equivalent to a random filter.
\end{proof}

As a consequence, after a sufficient number of elements, no filter can give a significant answer about the duplicate problem.

However, in most practical cases, it is not essential to detect duplicates since the start of the stream, but only duplicates on a sliding window, i.e., duplicates over the last $W$ elements for some integer $W$.

\subsection{Duplicate Detection over a Sliding Window}
In this section, we redefine the problem in order to include a sliding window.

\begin{definition}[Duplicate element, unseen element over a sliding window]
	Let $E = \{e_1, \dotsc, e_n, \dotsc\}$ be a stream over $\Gamma$, let $W\in \mathbb N$ be a sliding window size. $e_j$ is called a \emph{duplicate} in $E$ over the sliding window $W$ if and only if $\exists j - W \leq i < j, e_i = e_j$. Otherwise, $e_j$ is said to be \emph{unseen}.
\end{definition}

One observes that a duplicate or unseen element in $E = \{e_1, \dotsc, e_n\}$ over $W$ is equivalent to a duplicate or unseen element in $E' = \{e_{n-W}, \dotsc, e_n\}$ without a sliding window.

Similarly, we redefine the false positive, false negative, false positive rate and false negative rate.

\begin{definition}[False positive, false negative over a sliding window]
	Let $W$ be the size of a sliding window, $e$ be an unseen (resp. duplicate) element in $E = \{e_1, \dotsc, e_n\}$ over $W$, and let $\mathcal F$ be a filter. $e$ is a \emph{false positive} (resp. \emph{false negative}) over $W$ if $\mathsf{Detect}(e) = \DUPLICATE$ (resp. $\UNSEEN$).
\end{definition}

\begin{definition}[False positive rate, false negative rate over a sliding window]
	Let $E = \{e_1, \dotsc, e_n\}$ be a stream, $\mathcal F$ be a filter, $W$ be the size of a sliding window. The false positive (resp. false negative) rate of $\mathcal F$ over the first $i < n$ elements and the sliding window $W$, noted $\FPR^W_i$ (resp. $\FNR^W_i$), is defined as the number of false positives over $W$ divided by the number of unseen elements in $E$ over $W$ (resp. the number of false negative over $W$ divided by the number of duplicate elements over $W$).
\end{definition}

>>>>

\paragraph{Duplicate detection problem.}
We consider the following problem (duplicate detection problem, DPP): given a stream $S_t = \{s_t, s_{t-1}, \dotsc\}$ and an item $s^\star$, find whether $s^\star \in S_t$. It is understood that the stream $S_t$ is updated at every time increment. An \emph{efficient} solution to this problem provides an answer with bounded error $\epsilon$, using polynomially many resources (time and memory) in $1/\epsilon$.

@@

\begin{definition}[Duplicate and unseen element]
Let $E = \{e_i\}$ be a stream over $\Gamma$. Some element $e_j \in E$ is called a \emph{duplicate} if and only if there exists $i < j$ such that $e_i = e_j$. In that case we write $e_j \in^2 E$. Otherwise, $e_j$ is said to be \emph{unseen} and we write $e_j \notin^2 E$.
\end{definition}

@@

\begin{definition}[False positive, false negative]
	Let $e \in^2 E$ (resp. $e \notin^2 E$) and $\mathcal F$ be a filter. We say that $e$ is a \emph{false positive} (resp. \emph{false negative}) when $\mathcal F.\mathsf{Detect}(e) = \DUPLICATE$ (resp. $\UNSEEN$).
\end{definition}

@@

\subsection{Filter Saturation and Sliding Windows}
Unfortunately, duplicate detection over an infinite stream has the following effect. Since a filter can only store at most one element per bit \cite[Section 2.1]{10.1145/3297280.3297335}, a filter of size $M$ can information-theoretically remember at most $M$ elements. Thus after $fM$ insertions (with $f \gg 1$), at most a small fraction $1/f \ll 1$ of the stream can be stored, we say that the filter is \emph{saturated}.

In other terms, any filter is asymptotically as good as a random guess:
	\begin{theorem}[\cite{10.1145/3297280.3297335}]\label{thm:sat_u}
	Let $\FNR_\infty$ and $\FPR_\infty$ be the asymptotic false negative and false positive rates of a filter of $M$ bits of memory (i.e. the FNR and FPR on a stream of size going to infinity).

	If $|\Gamma| \gg M$, then $\FNR_\infty + \FPR_\infty = 1$, which characterises random filters (i.e. filters answering randomly to any query).
\end{theorem}
One way out of this situation is to lower our expectations and work on a sliding window:
\begin{definition}[Sliding window DDP]
	Given a stream of fixed size $S_t = \{s_t, \dotsc, s_{t-w}\}$ and an item $s^\star$, find whether $s^\star \in S_t$.
\end{definition} 
We therefore introduce new definitions suited to that setting:
\begin{definition}[Duplicate element, unseen element over a sliding window]
	Let $E = \{e_i\}$ be a stream over $\Gamma$, let $w\in \mathbb N$, an element $e_j \in E$ is called a \emph{duplicate} in $E$ over the sliding window of size $w$ if and only if there exists $j$ such that $j - w \leq i < j$  and $e_i = e_j$. In this case we write $e_j \in_w^2 E$. Otherwise, $e_j$ is said to be \emph{unseen} and we write $e_j \notin_w^2 E$.
\end{definition}
\begin{definition}[False positive, false negative over a sliding window]
	Let $w \in \mathbb N$, $e \notin_w^2 E$ (resp. $e \in_w^2 E$) and let $\mathcal F$ be a filter. We say that $e$ is a \emph{false positive} (resp. \emph{false negative}) over $w$ if $\mathsf{Detect}(e) = \DUPLICATE$ (resp. $\UNSEEN$).
\end{definition}

\begin{definition}[False positive rate, false negative rate over a sliding window]
	Let $E$ be a stream, $\mathcal F$ be a filter, $w \in \mathbb N$. The \emph{false positive} (resp. \emph{false negative}) rate of $\mathcal F$ over the first $i < n$ elements and the sliding window $w$, noted $\FPR^w_i$ (resp. $\FNR^w_i$), is defined as the number of false positives over $w$ divided by the number of unseen elements in $E$ over $w$ (resp. the number of false negative over $w$ divided by the number of duplicate elements over $w$).
\end{definition}
